---
output:
  rmarkdown::github_document
bibliography: "inst/REFERENCES.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# R/`uhalipw`

[![Travis-CI Build Status](https://travis-ci.org/nhejazi/uhalipw.svg?branch=master)](https://travis-ci.org/nhejazi/uhalipw)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/nhejazi/uhalipw?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/uhalipw)
[![Coverage Status](https://img.shields.io/codecov/c/github/nhejazi/uhalipw/master.svg)](https://codecov.io/github/nhejazi/uhalipw?branch=master)
[![CRAN](http://www.r-pkg.org/badges/version/uhalipw)](http://www.r-pkg.org/pkg/uhalipw)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/uhalipw)](https://CRAN.R-project.org/package=uhalipw)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)

> Nonparametric-Efficient IPW Estimation Based on the Undersmoothed Highly
> Adaptive Lasso

__Authors:__ [Nima Hejazi](https://nimahejazi.org) and [Mark van der
Laan](https://vanderlaan-lab.org/about/)

---

## What's `uhalipw`?

The `uhalipw` R package...

---

## Installation

Install the _most recent version_ from the `master` branch on GitHub via
[`remotes`](https://CRAN.R-project.org/package=remotes):

```{r gh-master-installation, eval = FALSE}
remotes::install_github("nhejazi/uhalipw")
```

Eventually, the package may make its way to [CRAN](https://CRAN.R-project.org).
At that point, a stable version may be installed via

```{r cran-stable-installation, eval = FALSE}
install.packages("uhalipw")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/nhejazi/uhalipw/issues).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/nhejazi/uhalipw/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `uhalipw` R package, please cite the following:

        @manual{hejazi2020uhalipw,
          author = {Hejazi, Nima S and {van der Laan}, Mark J},
          title = {{uhalipw}: Nonparametric-Efficient IPW Estimation Based on
            the Undersmoothed Highly Adaptive Lasso},
          year  = {2020},
          url = {https://github.com/nhejazi/uhalipw},
          note = {R package version 0.0.3}
        }

---

## License

&copy; 2019-2020 [Nima S. Hejazi](https://nimahejazi.org)

The contents of this repository are distributed under the MIT license. See below
for details:
```
MIT License

Copyright (c) 2019-2020 Nima S. Hejazi

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

---

## References

